---
title: "RMarkdown Introduction"
author: "Tatjana Kecojević"
date: "`r Sys.Date()`"
#output: html_document
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Markdown

This is my first R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

Material from today's workshop can be found on our [GitHub](tps://github.com/DataTeka/DSStoryWS) page.

When analysing data, a starting point is to examine the characteristics of each individual variable in the data set. The way to proceed depends upon the type of variable being examined. The variables can be one of two broad types:

1. **Attribute variable**: has its outcomes described in terms of its characteristics or attributes; 
2. **Measured variable**: has the resulting outcome expressed in numerical terms. 

## Including R Code

When you click the Knit button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r olympic}
olympic <- read.csv("data/athlete_events.csv")
olympic[1:3, ]
```

## Including Plots

You can also embed plots by setting `echo = FALSE` to the code chunk to prevent printing of the R code that generates the plot. For example:


```{r , echo=FALSE, warnings=FALSE, message=FALSE}
suppressPackageStartupMessages(library(dplyr))
library(ggplot2)
# get the data to be plotted
exyu <- olympic %>% 
  filter(Team %in% c("Bosnia and Herzegovina", "Croatia", "Serbia", "Serbia and Montenegro", "Montenegro", "Slovenia")) %>% 
  group_by(Team, Sex) %>% 
  summarize(total = n()) #<< 
exyu
# ggplot 
# we need a bar chart with each team on the x axis and number of male and female athlethes on the y axis.
ggplot(data = exyu, aes(x = Team, y = total, fill = Sex)) +
  geom_bar(stat="identity",  position="dodge", col = "black") +
# to make it read easier we will flip x & y coordinates
    coord_flip() +
# we will add description for x and y axies and title and subtitle  
    labs(x="ex YU country", y="No of athletes", 
      title = "Comparisons of M and F representatives in exYU Teams",
      subtitle = "for klikR workshop",
      caption = "Data from: kaggle - 120 years of Olympic history") +
# add the border on the graph  
    theme(panel.border = element_rect(fill = NA, colour = "black", size = 1)) +
#remove the grid lines
    theme(plot.title = element_text(size = 14, vjust = 2),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(), 
      axis.line = element_blank())
```

## Including Mathematical Equations

Let us fit the following model 
$$lifeExp = b_0 + b_1pop + b_2gdpPercap$$
which we write using the [LaTeX](https://en.wikibooks.org/wiki/LaTeX/Mathematics).
```{r}
m1 <- lm(olympic$Weight ~ olympic$Age)
summary(m1)
```

What do we think of this model?

#### **Let's discuss it next time we meet up.**
